### 2018-04-29 paper/intro.tex
halfway through intro, continuing. Added a ref for the CIA/NSA quote in adrien, and a macro for my comments.

% What's the problem?
Following the Weiserian vision, nowadays, we are
evolving in a new era of cooperation and interactions among humans.
The exponential proliferation of devices connected to the Internet is
drastically impacting human experience and enhancing human
interactions \cite{Dearman:2008, Oh:2017, Sohn:2008,Harper08}. It further opens new area of research such as
\textit{proxemics} applied to ubiquitous computing
\cite{Marquardt:2011}. Users are evolving in their social space with 
multiple kinds of computing devices that form a \textit{squad} of users'
devices, named \squad in the remainder of the
paper. Leveraging on the help of their \squad, users increase
their expectations in terms of connectivity and interactions
possibilities. For instance, as users bring
their devices close to one another, they expect to initiate cross
device interactions \textit{in situ} to share digital content
\cite{Oh:2017} among devices of their own \squad or with the ones 
from other users' \squad. More broadly, users expect to share digital content,
whatever the content location, from one
device to another in a seamless manner \cite{Dearman:2008}, as they
would have done in the old days with an USB key or with an email.

From a technical perspective, one major trend to provide a seamless
sharing of digital contents among users' devices is to leverage on the
cloud, which acts as a key enabler for ubiquitous computing. Users' data
are pushed online, into the cloud with either Dropbox, Google Docs,
OneDrive, Amazon Drive, etc. to make it accessible anytime,
anywhere to the users' devices.  In the last decades, a lot of research work has
emerged to promote the use of this offloading strategy to overcome scarce
resources of users' devices, and to provide an always-on availability
\cite{Zhang:2014,Gordon:2012,Chun:2011}. It paves the way of different kind
of clouds, such as mobile cloud computing, mobile-edge computing
and/or fog computing, that aim to use the nodes from the cloud
infrastructure that are the closest to end users
for offloading with an improved quality of service and user experience.

Hence, the cloud, and its variants constitute a perfect complement
to perform storage and computation on behalf of the resource
constrained devices. However, this cloud reliance comes at a strong price:
devices become totally tethered to a cloud ecosystem and provider. As
a direct consequence, users lose the control over their own data as it is
stored and spanned into multiple foreign data centers. Hence, it becomes difficult at 
the legal level to assess the privacy, and
confidentiality delivered to the end users. Further, at the practical
level, users' data can be harvested on their behalf, and without their
consent, by either the cloud provider or third parties.
Data sent by the user can be encrypted, but it is not enough
\cite{granick_2017, HooffLZZ15, HarnikPS10}.  As stated by a famous quotation from
NSA: "we kill people based on metadata''. Metadata such as the users'
location and activity can be still tracked, including the access
control history revealing what users share with whom and when.
Users’ devices become tracking and/or monitoring \squad that turn to
be the best spy over users.

>>>>

% What's the problem?
Following Mark Weiser's vision, we are nowadays
experiencing a new era of cooperation and interactions between humans.
The exponential proliferation of connected devices is
drastically impacting human experience and enhancing human
exchanges~\cite{Dearman:2008, Oh:2017, Sohn:2008,Harper08}. This evolution has opened new areas of research such as
\textit{proxemics} applied to ubiquitous computing
\cite{Marquardt:2011}. When evolving in their social space, users can today rely on
multiple kinds of computing devices that form a \textit{personal squad} of connected entities, or \squad as we shall call them in the remainder of the
paper. Helped by their \squad, users
have come to expect ever higher levels of connectivity and interaction
possibilities. For instance, as users bring
their devices close to one another, they expect to be able to initiate cross-device
interactions \textit{in situ} in order to share digital content
\cite{Oh:2017} among devices of their own \squad or with the ones 
from other users' \squad. More broadly, users expect to share digital content,
wherever the content, its sender, or its recipient might be located, from one
arbitrary device to another in a seamless manner~\cite{Dearman:2008}, as they
would have done in the old days with an USB key or with an email.

This thirst for seamless sharing explains the rise and ongoing popularity
of personal sharing platforms such as Dropbox, Google Docs,
OneDrive, Amazon Drive, etc. These solutions are able to share data on any device, between any user, almost
anytime, anywhere. They achieve this feat by relying heavily on
remotely hosted cloud services, which have now become a key enabler of ubiquitous ecosystems. 
% argument: cloud -> danger, 
%% From a technical perspective, one major trend to provide a seamless
%% sharing of digital contents among users' devices is to leverage on the
%% cloud, which acts as a key enabler for ubiquitous computing. Users' data
%% are pushed online, into the cloud with either Dropbox, Google Docs,
%% OneDrive, Amazon Drive, etc. to make it accessible anytime,
%% anywhere to the users' devices.
%% \commentFT{Removing the following, as seems to be a detour from main argument: \emph{In the last decades, a lot of research work has
%% emerged to promote the use of this offloading strategy to overcome scarce
%% resources of users' devices, and to provide an always-on availability
%% \cite{Zhang:2014,Gordon:2012,Chun:2011}. It paves the way of different kind
%% of clouds, such as mobile cloud computing, mobile-edge computing
%% and/or fog computing, that aim to use the nodes from the cloud
%% infrastructure that are the closest to end users
%% for offloading with an improved quality of service and user experience.}}
%% Hence, the cloud, and its variants constitute a perfect complement
%% to perform storage and computation on behalf of the resource
%% constrained devices.
Unfortunately, this reliance on the cloud comes at a strong price: 
%% users and their devices become totally tethered to the cloud ecosystem and its providers. As
%% a direct consequence,
users lose control over their own data, which become distributed over
multiple data centers, often in foreign jurisdictions, making it difficult for most users to grasp, review, and assess, the privacy, and
confidentiality guaranties they can expect. In practice, user data hosted in the cloud is routinely harvested, often without explicit
consent, either by cloud providers or third parties.

Although data encryption is often proposed as a first line of defense for cloud-hosted data, it is unfortunately  insufficient, because it does not protect meta-data information~\cite{granick_2017, HooffLZZ15, HarnikPS10}. This turns out to be a critical limitation, as famously emphasized by a former NSA and CIA director: "We kill people based on metadata.''~\cite{NaughtonTheGuardian2016} Metadata such as a user'
location and activity can be still tracked, including the access
control history revealing what each user shared with whom and when.
User devices become tracking and/or monitoring \squad that turn to
be the best spy over users.
